---
title: "Online News Popularity - Analysis"
author: "cbcoursera1"
date: "Monday, June 29, 2015"
output: html_document
---

```{r}

# Code allowing for figures dynamically in Appendix

library(knitr)
opts_chunk$set(echo = F, warnings = F, message = F, comment = F, include = T, cache = T)

```

```{r}

# Report format

# indicates a comment (ie code section header)
## indicates a subcomment (comments within code section )
### indicates a commented piece of code - eg a function you tried but doesn't contribute further down the line

# During exploratory analysis let's keep our code chunks separate

# Feel free to add some formatting advice here

```

```{r, 1}

# Libraries

library(data.table)
library(ggplot2)

# Ingest the data

download.file("http://archive.ics.uci.edu/ml/machine-learning-databases/00332/OnlineNewsPopularity.zip", "OnlineNewsPopularity.zip")
unzip("OnlineNewsPopularity.zip")
onp <- fread("OnlineNewsPopularity/OnlineNewsPopularity.csv", sep = ", ", header = T, data.table = T)
onp.bak <- onp # backup
onp <- onp.bak # revert from backup
### lapply(mean, onp) # quick visual check that means match summary provided with data

# Clean data

setnames(onp, gsub("_", ".", names(onp))) # fix column titles
setnames(onp, gsub(" ", "", names(onp))) # fix column titles
### lapply(onp, class) # look at the classes, mostly they look fine
### onp[, grep(".is.", names(onp))] <- lapply(onp[, grep(".is.", names(onp)), with = F], as.logical)

```

=========================================================================

# Everything after here is WIP

=========================================================================

```{r}

## Let's do three things with the data

## turn the weekday.is.monday etc variables into a factor variable of dayofweek

onp[, day.of.week := factor(rep("", nrow(onp)), levels = c("mon", "tue", "wed", "thu", "fri", "sat", "sun"))]  ## create empty factor variable

dow <- data.table("s" = c("mon", "tue", "wed", "thu", "fri", "sat", "sun"), "l" = c("monday", "tuesday", "wednesday", "thursday", "friday", "saturday", "sunday")) # create data table of short and long day names

## CB ===================================================================
## Having a hard time with this. My goal is to binary search the data.table in
## each weekday.is column, and if it's a 1 (ie TRUE) then set the day.of.week
## column with the corresponding factor
## My plan was to iterate through the for statement for each pair, eg "mon", "monday"
## The problem right now is that I can't get a simple binary search to work
## on the data.table, eg    
##
## onp[weekday.is.monday == 1]
##
## it tells me the object is not found, but it is the column name. Removing the
## .s also does not help. However
##
## onp[url == 1]
##
## works fine. Maybe because it's the first column? Anyway, help appreciated.

for (i in nrow(dow)) {
    
    onp[paste0("weekday.is.", dow[i, l]) == T, day.of.week := dow[i, s]]

}   ## iterate across days of week filling the day.of.week column

## turn the data.channel.is.bus etc variables into a factor varaible of channel (same method as dow)

## grab the date from the URL


```

# Appendix

```{r}

opts_chunk$set(include = F)

```

### Source

UCI Machine Learning Repository - Online News Popularity Data Set, donated 5/31/15
https://archive.ics.uci.edu/ml/datasets/Online+News+Popularity

### Citation

    K. Fernandes, P. Vinagre and P. Cortez. A Proactive Intelligent Decision
    Support System for Predicting the Popularity of Online News. Proceedings
    of the 17th EPIA 2015 - Portuguese Conference on Artificial Intelligence,
    September, Coimbra, Portugal.